---
title: "STAT6348 HW1"
author: "Minzhe Zhang mxz163730"
date: "2017/10/6"
output: html_document
---

### 1. Question 1
Based on the individual ANOVA tests below, thickness and overrun are not the same among different housing conditions. From the manova test, we know there is at least one housing condition has different characteristics from the rest.
```{r}
### read and normalize data
egg.data <- read.csv("data.csv", stringsAsFactors = FALSE)
egg.data[,-1] <- apply(egg.data[,-1], 2, scale)
### fit manova model
fit <- manova(as.matrix(egg.data[,-1]) ~ HOUSING, data = egg.data)
### individual anova test
summary.aov(fit)
### manova test
summary(fit, test = "Wilks")
```

### 2. Question 2
For every two housing conditions, we do the Hotellings T2 test. From the test result below, we see that housing condition cage has significant characteristics with other conditions, while the rest don't have significant differene with each other.
```{r}
suppressMessages(library(ICSNP))
egg.cage <- egg.data[egg.data$HOUSING == "CAGE",-1]
egg.free <- egg.data[egg.data$HOUSING == "FREE",-1]
egg.barn <- egg.data[egg.data$HOUSING == "BARN",-1]
egg.organic <- egg.data[egg.data$HOUSING == "ORGANIC",-1]
HotellingsT2(egg.cage, egg.free, test = "chi")
HotellingsT2(egg.cage, egg.barn, test = "chi")
HotellingsT2(egg.cage, egg.organic, test = "chi")
HotellingsT2(egg.free, egg.barn, test = "chi")
HotellingsT2(egg.free, egg.organic, test = "chi")
HotellingsT2(egg.barn, egg.organic, test = "chi")
```

### 3. Question 3
Mardia's multivariate normality test show all four groups follow multivariate normal distribution.
```{r}
suppressMessages(library(MVN))
mardiaTest(egg.cage)
mardiaTest(egg.free)
mardiaTest(egg.barn)
mardiaTest(egg.organic)
```

### 4. Question 4
From the above statistical test, we see that cage egg is significantly from the egg with other housing conditions. Several study also show that cage raised hen laid egg has thinner shell than barn and aviary[1-3]. These may due to more daily feed intake and more locomotion activity of hen in a more free environment.


#### Referrence
1. Pavlovski, Zlatica, S. Hopic, and M. Lukic. "Housing systems for layers and egg quality." Biotechnology in Animal Husbandry (2001).
2. Hidalgo, A., et al. "A market study on the quality characteristics of eggs from different housing systems." Food Chemistry 106.3 (2008): 1031-1038.
3. Ahammed, M., et al. "Comparison of aviary, barn and conventional cage raising of chickens on laying performance and egg quality." Asian-Australasian journal of animal sciences 27.8 (2014): 1196.