---
title: "Final"
author: "Minzhe Zhang"
date: "2017/12/10"
output: html_document
---

### Question 1
From the result below, we see that LD1 and LD2 achieve a good separation of the response variable. In LD1 and LD2, variable **Mexico**, **Jalisco**, **Nuevo**, **Tamaulipas** and **Nayarit** are important ones to explain dynamics of chikungunaya.
```{r}
### read data
library(MASS)
gdt.data <- read.csv("FinalExam_GDT_Chikungunya2017.csv")
# gdt.data[,-c(1,2)] <- apply(gdt.data[,-c(1,2)], 2, scale)
gdt.lda <- lda(gdt.data$TotalCumulativeChikungunayaCases ~ gdt.data$Mexico + gdt.data$BajaCalifornia + gdt.data$DistritoFederal + gdt.data$Jalisco + gdt.data$Estado + gdt.data$Morelos + gdt.data$Nuevo + gdt.data$Oaxaca + gdt.data$Quintana.Roo + gdt.data$Sonora + gdt.data$Tamaulipas + gdt.data$Veracruz + gdt.data$Yucatan + gdt.data$Chiapas + gdt.data$Colima + gdt.data$Nayarit + gdt.data$Sinaloa + gdt.data$Tabasco)
gdt.lda
```

### Question 2
From the hierarchical clustering plot below, using 0.5 as cut off, the following variables could be clustered together.

- **BajaCalifornia** and **Sonora**.
- **Jalisco**, **Tabasco**, **Nuevo** and **Nayarit**.
- **Yucatan**, **Tamaulipas** and **Veracruz**.
- **DistritoFederal** and **Chiapas**.
- **Quintana Roo** and **Colima**.
```{r}
clust <- hclust(dist(t(gdt.data[,-c(1,2)])))
plot(clust)
abline(h = 0.5, col = "red")
```