---
title: "Midterm"
author: "Minzhe Zhang mxz163730"
date: "2017/11/3"
output: html_document
---

### Question 1
The computed principal components are listed below.
```{r}
suppressMessages(library(nFactors))
### read and normalize data
air.data <- read.csv("data.csv")
air.data <- apply(air.data, 2, scale)
### pca
air.pca <- prcomp(air.data)
print(air.pca)
print(summary(air.pca))
```

### Question 2
From the Bartlett, Anderson and Lawley Procedures, 6 out of 7 components are significant.
```{r}
### Determine the significant PCA
pca.test <- nBartlett((air.pca$sdev)^2, N = dim(air.data)[1], alpha = 0.05, details = TRUE)
summary(pca.test)
```

### Question 3
The sample covariance matrix is the following
```{r}
new.data <- air.data[,c(1,2,5,6)]
cor.mat <- cor(new.data)
cor.mat
```
#### (a) Principal component with 1 and 2 factors
```{r}
print(principal(new.data, nfactors = 1))
print(principal(new.data, nfactors = 2))
```

#### (b) Varimax rotation of both m = 1 and m = 2 solutions. 
Solar radiation and O3 contribute mostly to the only factor when m = 1. When m = 2, Solar radiation and O3 are still the main components of factor 1; wind and O3 are the main component of factor 2.
```{r}
FA1 <- fa(new.data, nfactors = 1, rotate = "varimax")
FA2 <- fa(new.data, nfactors = 2, rotate = "varimax")
print(FA1)
print(FA2)
```

#### (c) Discussion
1. Two factor is enough to explain almost all the varibility in the data. 
2. The result of prinicpal component analysis anf factor analysis is close but not exactly same, indicating there is similarity and distinction in these two type of methods.